Blame and Coercion: Together Again for the Second Time

### 2019-10-16 paper.tex

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. -Jeremy}

@@

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove their \lazyUD{} coercion simulates the Blame Calculus in 
\cite{wadler2009well}. It is unknown, however, whether the simulation in the 
other direction also works.

@@

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bisimulate the 
corresponding Blame Calculi.

\citet{garcia2013calculating} develops
threesome for other blame strategies. After studying in depth the connection 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is mostly affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the coercion calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} coercion in \citet{siek2012interpretations} and prove that 
their revised coercion calculus bisimulates the \lazyUD{} Blame Calculus.

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
type constructor at the top. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame. Shallow-consistency is reflexive, symmetric, but 
not transitive.

Let $ e $ ranges over terms, including all terms from $ \mathtt{STLC+} $ and, 
in addition, casts and blames. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly. 
The end of the figure shows the typing rules for the additional terms.

Let $ o $ ranges over observations. They are what would be observed if a 
program terminates. Observations include all value constructors and blames.

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}).
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hypercoercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hypercoercions.

@@

Let $ c $ ranges over casts. A cast is a triple of a type, a label, and a type.
Let $ v $ ranges over values. A value is either a function, the 
element of unit, a pair, a left injection, a right injection, or a casted value.
The cast around a casted value is not arbitrary, as shown by value typing: the 
source type must be a pre-type, and the source and the target must be 
shallow-consistent.
Let $ r $ ranges over cast results. A cast result is either a success, which 
brings a value, or a failure, which brings a blame label.
Let $ s $ ranges over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.
Let $ \kappa $ ranges over continuations. $ \mathtt{mt} $ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

In the reduction rules, to evaluate a cast expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one. Along the way, the machine accumulate the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

@@

We use bisimulation to proof equivalence between the two machines.
The state bisimilarity ($\eqvDS{C}{s}{s}$) is mostly derived from the 
definitions of $ \mathcal{D} $. Exceptions include the bisimilarity relations 
between continuations, between values, and from continuations to 
pre-continuations. We omit subscripts when there is no ambiguity.

@@

\begin{lemma}[Weak Bi-simulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeDBreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeDBreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeDBreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

>>>>

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. The rules are tricky...
  some words don't have a plural form, like ``blame''. -Jeremy}

\todo[inline]{The first paragraph of the introduction should summarize
  the entire introduction, and in some sense, the entire paper. -Jeremy}

\todo[inline]{BTW, using the Coercion and Blame paper as a model for
  how to write papers is perhaps not the best idea. That paper intentionally
  diverged from the normal paper format in several ways. You can
  do that too after you've become a master at writing papers.
  For now you are much better off following the advice of
  Simon Peyton Jones about how to write a paper (google it).}

@@

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove their \lazyUD{} coercion simulates the Blame Calculus in 
\cite{wadler2009well}. It is unknown, however, whether the simulation in the 
other direction also works.
\todo[inline]{Is it important to this paper whether there is a
  simulation in the other direction? -Jeremy}

@@

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
\todo[inline]{Not ``first''... lazy D is in the 2009 paper, right?
  Or did you mean something else? -Jeremy}
\todo[inline]{The ``Blame Calculus'' is a \lazyUD{} cast calculus.
  There is no such thing as a \lazyD{} Blame Calculus. -Jeremy}
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bisimulate the 
corresponding Blame Calculi.

\citet{garcia2013calculating} develops
threesome for other blame strategies. After studying in depth the connection 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is mostly affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the coercion calculi of
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

\citet{siek2015blame} revisit the coercion-based approach. They revise
the \lazyUD{} coercion in \citet{siek2012interpretations} and prove
that their revised coercion calculus bisimulates the \lazyUD{} Blame
Calculus \citet{wadler2009well}.

@@

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
type constructor at the top. Shallow-inconsistency is the root of all blame in 
lazy cast strategies -- casting a value to a shallowly inconsistent type leads 
to a blame. Shallow-consistency is reflexive, symmetric, but 
not transitive.

Let $ e $ ranges over terms, including all terms from $ \mathtt{STLC+} $ and, 
in addition, casts and blame. Unlike $ \mathtt{STLC+} $, we annotate the 
codomain of lambda abstractions explicitly.
\todo[inline]{Why do you annotate the codomain of lambda abstractions
  explicitly? -Jeremy}
The end of \figref{fig:blame-static} shows the typing rules for the
additional terms.

Let $ o $ ranges over observations. They are what would be observed if a 
program terminates. Observations include all value constructors and blames.
\todo[inline]{``blames'' is incorrect. See
https://www.wordhippo.com/what-is/the-plural-of/blame.html}

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines \citep{felleisen1986control}.
\todo[inline]{Briefly present the main ideas regarding how a CEK
  machine works. (2-3 sentences) -Jeremy}
\todo[inline]{Use citep to put a citation in parentheses, as I've
  done above. -Jeremy }
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hypercoercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hypercoercions.

@@

Let $ c $ ranges over casts. A cast is a triple of a type, a label, and a type.
\todo[inline]{``ranges'' should be ``range'', not plural. -Jeremy}
Let $ v $ ranges over values. A value is either a function, the 
element of unit, a pair, a left injection, a right injection, or a casted value.
The cast around a casted value is not arbitrary, as shown by value typing: the 
source type must be a pre-type, and the source and the target must be 
shallow-consistent.
Let $ r $ ranges over cast results. A cast result is either a success, which 
brings a value, or a failure, which brings a blame label.
Let $ s $ ranges over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.
Let $ \kappa $ ranges over continuations. $ \mathtt{mt} $ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

In the reduction rules, to evaluate a cast expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one.
\todo[inline]{Change the rules for case to be small step, i.e.,
 the case function should not be recursive. -Jeremy}
Along the way, the machine accumulate the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

@@

%%<<<<<<< HEAD
With the properties of surely \lazyD{} CR, it is straightforward to proof that 
all machines instantiated with these CRs bisimulate each other. 
The state bisimilarity ($\eqvS{C}{C}{s}{s}$) is 
derived from the definition of states.
\todo[inline]{The term ``bisimilarity'' has a different technical meaning from
  ``bisimulation relation''. I don't think you are using it properly
  in the above sentence. -Jeremy}
We omit subscripts when there is no 
ambiguity.
\begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}

\begin{proposition}[Strong bisimulation among $ \mathcal{S}(\cdot) $]
	\label{thm:CEKS-bisim}
	If 
	$ C_1 $ and $ C_2 $ are surely \lazyD,
	$ s_1, s_2 \in S(C_1) $ and $ s_3 \in S(C_2) $,
	$ s_1 \approx s_3 $,
	$ \judgeSreduce{C_1}{s_1}{s_2} $,
	then there exists an $ s_4 $ such that
	\begin{itemize}
	\item $ s_2 \approx s_4 $ and
	\item $ s_3 \longrightarrow_{S(C_2)} s_4 $
	\end{itemize}
\end{proposition}
\begin{proof}
	The key ideas of this proof are undoing sequencing with the property (2) of 
	surely \lazyD{} CR, and handling all possibly uses of $ 
	cast(T,l,T) $ with property (3)-(11).
\end{proof}

\begin{proposition}[Equivalence of Surely \lazyD{} Cast Representations]
	\label{thm:surely-lazyD-eqv}
	If $ \judgetype{\emptyset}{e}{T} $, $ o : T $, and $ C_1 $ and $ C_2 $ 
	are surely \lazyD cast representations,
	\[
	e \Downarrow_{S(C_1)} o \; \text{if and only if} \; 
	e \Downarrow_{S(C_2)} o
	\]
\end{proposition}
\begin{proof}
	Immediately from Proposition~\ref{thm:CEKS-bisim}.
\end{proof}

\subsection{Relating All Surely \lazyD{} Cast Representations to $\mathcal{D}$}

So far, we have related all surely \lazyD{} CRs, i.e. they are equivalent to 
each other. But this does not mean they respect the semantics of the \lazyD{} 
Blame Calculus. We solve prove this part in this subsection.

The state bisimilarity here ($\eqvDS{C}{s}{s}$) is mostly derived from the 
%%=======
We use bisimulation to proof equivalence between the two machines.
The state bisimilarity ($\eqvDS{C}{s}{s}$) is mostly derived from the 
%%>>>>>>> 6d40f5173c83f5f7f7b37a4128391ca516854ac9
definitions of $ \mathcal{D} $. Exceptions include the bisimilarity relations 
between continuations, between values, and from continuations to 
pre-continuations. We omit subscripts when there is no ambiguity.

@@

%%<<<<<<< HEAD
\begin{lemma}[Weak bisimulation between $ \mathcal{S}(\cdot) $ and $ 
\mathcal{D} $]
\label{thm:surely-monoidic-reduce}
For all cast representations $ C $, 
if $ C $ is surely \lazyD{} and monoidic,
$ s_1 \in \mathcal{D} $,
$ s_2 \in \mathcal{S}(C) $, and
$ s_1 \approx s_2 $, 
\begin{enumerate}
	\item If $ \judgeDBreduce{s_1}{s_3} $,
	then
	there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
	such that \begin{itemize}
		\item $ s_5 \approx s_4 $, and
		\item $ \judgeSreduce{C}{s_2}{s_4} $,
		\item $ \judgeDBreduceTrans{s_3}{s_5} $
	\end{itemize}
	\item If $ \judgeSreduce{C}{s_2}{s_4} $,
	then
	there exist $ s_3 \in \mathcal{D} $,
	such that \begin{itemize}
		\item $ s_3 \approx s_4 $
		\item $ \judgeDBreduceTrans{s_1}{s_3} $
	\end{itemize}
\end{enumerate}
%%=======
\begin{lemma}[Weak Bi-simulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeDBreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeDBreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeDBreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
%%>>>>>>> 6d40f5173c83f5f7f7b37a4128391ca516854ac9
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}
